---
title: "smFISH quantification report `r format(Sys.Date(), '%Y-%m-%d')`"
author: "Naly Torres"
date: "1/30/2024"
output: html_document
---


```{r echo=FALSE, warning=FALSE} 
library(png)
library(png)
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)

```


Violin Plots

```{r}
# Function to read CSV, replace Image ID with CSV name, and modify column names
read_modify_csv <- function(file_path, subdirectory) {
  my_df <- read.csv(file_path)
  my_df$Image_ID <- basename(file_path)
  my_df$subdirectory <- subdirectory
  my_df <- my_df[, colnames(my_df) != "X"]
  current_colnames <- colnames(my_df)
  new_colnames <- gsub("\\.", " ", current_colnames)
  colnames(my_df) <- new_colnames
  return(my_df)
}

# Parent directory containing CSV files for all conditions
parent_directory <- "/Users/nalytorres/Documents/01_bigfish/02_data/DG-quantification"

# List all subdirectories (conditions) in the parent directory
subdirectories <- list.dirs(parent_directory, full.names = FALSE, recursive = FALSE)

# Initialize an empty list to store data frames
all_data <- list()

# Loop through subdirectories (conditions)
for (subdirectory in subdirectories) {
  # List all CSV files in the current subdirectory
  csv_files <- list.files(path = file.path(parent_directory, subdirectory), pattern = "*.csv", full.names = TRUE)
  
  # Loop through CSV files, read data, replace Image ID, and modify column names
  for (file_path in csv_files) {
    current_data <- read_modify_csv(file_path, subdirectory)
    all_data[[file_path]] <- current_data
  }
}

# Combine all data frames into one data frame
combined_data <- bind_rows(all_data)
combined_data

```
```{r}
# Find the row indices where either condition is met
rows_to_remove <- which(combined_data$`lin 41 mRNA clusters` > 50 | combined_data$`lin 41 mRNA molecules` < 100)

# Remove the rows
combined_data <- combined_data[-rows_to_remove, ]

# Print the updated data
print(combined_data)

```


```{r}
# Load required libraries
library(dplyr)

# Extract cell stage and strain from the subdirectory column
combined_data <- combined_data %>%
  mutate(Cell_Stage = gsub("_.*", "", subdirectory),
         Strain = gsub("^(2-cell_|4-cell_)", "", subdirectory))

# Check the first few rows to verify the extraction
tail(combined_data)


str(combined_data)

# Save combined_data as a CSV file
write.csv(combined_data, file = "lin-41-mRNA-abundance.csv", row.names = FALSE)


```






```{r}
# Calculate the mean of each group
mean_values <- aggregate(cbind(`set 3 mRNA molecules`, `lin 41 mRNA molecules`) ~ subdirectory, data = combined_data, FUN = mean)

# View the mean values
print(mean_values)

# Assuming your data frame is named 'df'
# Separate the 'subdirectory' column into 'Strain' and 'Cell_Stage'
df <- separate(mean_values, subdirectory, into = c("Cell_Stage", "Strain"), sep = "_", remove = FALSE)

# Reorder the columns if needed
df <- df[, c("Cell_Stage", "Strain", "set 3 mRNA molecules", "lin 41 mRNA molecules")]

# Print the updated data frame
print(df)

str(df)

```




```{r}
# Create the scatter plot with colored points and lines connecting 2-cell and 4-cell of each strain
p <- ggplot(df, aes(x = Cell_Stage, y = `lin 41 mRNA molecules`, color = Strain, group = Strain)) +
  geom_point(position = position_dodge(width = 0.5)) +  # Adjust dodge width as needed
  geom_line(position = position_dodge(width = 0.5), color = "grey") +  # Adjust dodge width and line color as needed
  scale_x_discrete(limits = c("2-cell", "4-cell")) +
  scale_color_manual(values = strain_colors) +
  labs(x = "Cell Stage", y = "lin-41 mRNA Abundance", color = "Strain") +
  ggtitle("lin-41 mRNA Abundance at Different Cell Stages by Strain") +
  theme_minimal() +  # Use a minimal theme
  theme(panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

# Print the scatter plot
print(p)

# Save the plot as SVG with specified output path
ggsave("/Users/nalytorres/Documents/01_bigfish/02_data/DG-quantification/slope-lin-41-plot.svg", p, width = 10, height = 6)



```


```{r}

# Define the custom order for subdirectories
custom_order <- c("2-cell_N2", "2-cell_DG3913", "2-cell_DG5410", "2-cell_DG5329", "2-cell_DG5399", "2-cell_DG5398", "2-cell_DG5779","4-cell_N2","4-cell_DG3913", "4-cell_DG5410","4-cell_DG5329", "4-cell_DG5399", "4-cell_DG5398","4-cell_DG5779")  # Add your actual order

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)


# Create a combined violin and boxplot (flipped vertically)
p <- ggplot(combined_data, aes(x = subdirectory, y = `lin 41 mRNA molecules`, fill = subdirectory)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 0.99), outlier.shape = NA) +
  geom_violin(alpha = 0.7, color = "lightblue", size = 0.5, position = position_dodge(width = 0.99)) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 0.5) +
  geom_vline(xintercept = c(3150, 1100), color = c("black", "red"), linetype = c("dashed", "longdash"), alpha = 0.5, size = 1) +  # Add two vertical lines
  labs(title = "Abundance of lin-41 mRNA in 3'UTR deletion mutants",
       x = "3'UTR deletion strain",
       y = "lin 41 mRNA abundance \n (total molecules)",
       fill = "Cell stage") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("2-cell_N2", "4-cell_DG5410"),
                    labels = c("2-cell", "4-cell")) +
  scale_x_discrete(labels = c("2-cell_N2" = "N2 2-cell ",
                              "4-cell_N2" = "N2 4-cell ",
                              "2-cell_DG5410" = "DG5410 2-cell ",
                              "4-cell_DG5410" = "DG5410 4-cell ",
                              "2-cell_DG5398" = "DG5398 2-cell ",
                              "4-cell_DG5398" = "DG5398 4-cell ",
                              "2-cell_DG5779" = "DG5779 2-cell ",
                              "4-cell_DG5779" = "DG5779 4-cell ",
                              "2-cell_DG5399" = "DG5399 2-cell ",
                              "4-cell_DG5399" = "DG5399 4-cell ",
                              "2-cell_DG5329" = "DG5329 2-cell ",
                              "4-cell_DG5329" = "DG5329 4-cell ")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate labels on x-axis
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(ylim = c(0, 12000)) +
  guides(fill = guide_legend(nrow = 2))  # Limit the legend to two rows

# Display the plot
print(p)


# Save the plot as SVG with specified output path
ggsave("/Users/nalytorres/Documents/01_bigfish/02_data/DG-quantification/lin-41-side-by-side.svg", p, width = 10, height = 6)

```



```{r}
# Define the custom order for subdirectories
custom_order <- c("2-cell_N2", "2-cell_DG3913", "2-cell_DG5410", "2-cell_DG5329", "2-cell_DG5399", "2-cell_DG5398", "2-cell_DG5779","4-cell_N2","4-cell_DG3913", "4-cell_DG5410","4-cell_DG5329", "4-cell_DG5399", "4-cell_DG5398","4-cell_DG5779")  # Add your actual order

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Create a combined violin and boxplot (flipped vertically)
p <- ggplot(combined_data, aes(x = factor(subdirectory, levels = custom_order), y = `lin 41 mRNA molecules`)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 10), outlier.shape = NA) +
  geom_violin(alpha = 0.6, fill = "gray", size = 0.4, width = 6, position = position_dodge(width = 10)) + 
  geom_jitter(height = 0, width = 0.2, color = "black", size = 0.5) +
  labs(title = "Abundance of lin-41 mRNA in 3'UTR deletion mutants",
       x = "3'UTR deletion strain",
       y = "lin 41 mRNA abundance \n (total molecules)") +  # Remove fill legend
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate labels on x-axis
        legend.position = "none",  # Remove legend
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.text = element_blank()) +
  coord_cartesian(ylim = c(0, 12000))

# Display the plot
print(p)

# Save the plot as SVG with specified output path
ggsave("/Users/nalytorres/Documents/01_bigfish/02_data/DG-quantification/lin-41-side-by-side.svg", p, width = 10, height = 6)


```

```{r}
# Define the custom order for subdirectories
custom_order <- c("2-cell_N2", "2-cell_DG3913", "2-cell_DG5410", "2-cell_DG5329", "2-cell_DG5399", "2-cell_DG5398", "2-cell_DG5779","4-cell_N2","4-cell_DG3913", "4-cell_DG5410","4-cell_DG5329", "4-cell_DG5399", "4-cell_DG5398","4-cell_DG5779")  # Add your actual order

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Create a combined violin and boxplot (flipped vertically)
p <- ggplot(combined_data, aes(x = factor(subdirectory, levels = custom_order), y = `lin 41 mRNA molecules`)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 1.2), outlier.shape = NA) +
  geom_violin(alpha = 0.7, fill = "gray", size = 0.5, position = position_dodge(width = 5)) + 
  geom_jitter(width = 0.2, color = "black", size = 0.5) + 
  labs(title = "Abundance of lin-41 mRNA in 3'UTR deletion mutants",
       x = "3'UTR deletion strain",
       y = "lin 41 mRNA abundance \n (total molecules)") +  # Remove fill legend
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate labels on x-axis
        legend.position = "none",  # Remove legend
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.text = element_blank()) +
  coord_cartesian(ylim = c(0, 12000))

# Display the plot
print(p)

```




```{r}
# Define the custom order for subdirectories
custom_order <- c("2-cell_N2", "2-cell_DG3913", "2-cell_DG5410", "2-cell_DG5329", "2-cell_DG5399", "2-cell_DG5398", "2-cell_DG5779","4-cell_N2","4-cell_DG3913", "4-cell_DG5410","4-cell_DG5329", "4-cell_DG5399", "4-cell_DG5398","4-cell_DG5779")  # Add your actual order

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Create a combined violin and boxplot (flipped vertically)
p <- ggplot(combined_data, aes(x = subdirectory, y = `lin 41 mRNA molecules`, fill = subdirectory)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 0.99), outlier.shape = NA, color = "gray") +
  geom_violin(alpha = 0.7, color = "gray", size = 0.5, position = position_dodge(width = 0.99)) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 0.5) +
  geom_vline(xintercept = c(3150, 1100), color = c("black", "red"), linetype = c("dashed", "longdash"), alpha = 0.5, size = 1) +  # Add two vertical lines
  labs(title = "Abundance of lin-41 mRNA in 3'UTR deletion mutants",
       x = "3'UTR deletion strain",
       y = "lin 41 mRNA abundance \n (total molecules)",
       fill = "Cell stage") +  # Set the legend title here
  scale_fill_manual(values = c("gray"),  # Set fill color to gray
                    breaks = c("2-cell_N2", "4-cell_DG5410"),
                    labels = c("2-cell", "4-cell")) +
  scale_x_discrete(labels = c("2-cell_N2" = "N2 2-cell ",
                              "4-cell_N2" = "N2 4-cell ",
                              "2-cell_DG5410" = "DG5410 2-cell ",
                              "4-cell_DG5410" = "DG5410 4-cell ",
                              "2-cell_DG5398" = "DG5398 2-cell ",
                              "4-cell_DG5398" = "DG5398 4-cell ",
                              "2-cell_DG5779" = "DG5779 2-cell ",
                              "4-cell_DG5779" = "DG5779 4-cell ",
                              "2-cell_DG5399" = "DG5399 2-cell ",
                              "4-cell_DG5399" = "DG5399 4-cell ",
                              "2-cell_DG5329" = "DG5329 2-cell ",
                              "4-cell_DG5329" = "DG5329 4-cell ")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate labels on x-axis
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(ylim = c(0, 12000)) +
  guides(fill = guide_legend(nrow = 2))  # Limit the legend to two rows

# Display the plot
print(p)

```

```{r}
# Define the custom order for subdirectories
custom_order <- c("2-cell_N2", "2-cell_DG3913", "2-cell_DG5410", "2-cell_DG5329", "2-cell_DG5399", "2-cell_DG5398", "2-cell_DG5779","4-cell_N2","4-cell_DG3913", "4-cell_DG5410","4-cell_DG5329", "4-cell_DG5399", "4-cell_DG5398","4-cell_DG5779")  # Add your actual order

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Create a combined violin and boxplot (flipped vertically)
p <- ggplot(combined_data, aes(x = factor(subdirectory, levels = custom_order), y = `lin 41 mRNA molecules`)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 10), outlier.shape = NA) +
  geom_violin(alpha = 0.6, fill = "gray", size = 0.4, width = 3, position = position_dodge(width = 10)) + 
  geom_jitter(aes(color = subdirectory), height = 0, width = 0.2, size = 0.5) + # Map color to subdirectory variable
  labs(title = "Abundance of lin-41 mRNA in 3'UTR deletion mutants",
       x = "3'UTR deletion strain",
       y = "lin 41 mRNA abundance \n (total molecules)") +  # Remove fill legend
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),  # Rotate labels on x-axis
        legend.position = "none",  # Remove legend
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.text = element_blank()) +
  coord_cartesian(ylim = c(0, 12000))

# Display the plot
print(p)

# Save the plot as SVG with specified output path
ggsave("/Users/nalytorres/Documents/01_bigfish/02_data/DG-quantification/lin-41-side-by-side-colors.svg", p, width = 10, height = 6)


```


```{r}

# Define the custom order for subdirectories
custom_order <- c("4-cell_DG5399", "2-cell_DG5399","4-cell_DG5398", "2-cell_DG5398","4-cell_DG5779", "2-cell_DG5779","4-cell_DG5410","2-cell_DG5410", "4-cell_DG5329","2-cell_DG5329","4-cell_DG3913", "2-cell_DG3913", "4-cell_N2", "2-cell_N2" )  # Add your actual order

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)


# Create a combined violin and boxplot (flipped vertically)
p <- ggplot(combined_data, aes(x = `lin 41 mRNA molecules`, y = subdirectory, fill = subdirectory)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 0.99), outlier.shape = NA) +
  geom_violin(alpha = 0.7, color = "white", size = 0.5, position = position_dodge(width = 0.99)) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 0.5) +
  geom_vline(xintercept = c(3150, 1100), color = c("black", "red"), linetype = c("dashed", "longdash"), alpha = 0.5, size = 1) +  # Add two vertical lines
  labs(title = "Abundance of lin-41 mRNA in 3'UTR deletion mutants",
       y = "3'UTR deletion strain",
       x = "lin 41 mRNA abundance \n (total molecules)",
       fill = "Cell stage") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("2-cell_N2", "4-cell_DG5410"),
                    labels = c("2-cell", "4-cell")) +
  scale_y_discrete(labels = c("2-cell_N2" = "N2 2-cell ",
                              "4-cell_N2" = "N2 4-cell ",
                              "2-cell_DG5410" = "DG5410 2-cell ",
                              "4-cell_DG5410" = "DG5410 4-cell ",
                              "2-cell_DG5398" = "DG5398 2-cell ",
                              "4-cell_DG5398" = "DG5398 4-cell ",
                              "2-cell_DG5779" = "DG5779 2-cell ",
                              "4-cell_DG5779" = "DG5779 4-cell ",
                              "2-cell_DG5399" = "DG5399 2-cell ",
                              "4-cell_DG5399" = "DG5399 4-cell ",
                              "2-cell_DG5329" = "DG5329 2-cell ",
                              "4-cell_DG5329" = "DG5329 4-cell ")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(angle = 0, hjust = 1),  # Rotate labels on y-axis
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(xlim = c(0, 12000)) +
  guides(fill = guide_legend(nrow = 2))  # Limit the legend to two rows

# Display the plot
print(p)

```

```{r}

# Save the plot as SVG with specified output path
ggsave("/Users/nalytorres/Documents/01_bigfish/02_data/DG-quantification/lin-41-plot.svg", p, width = 10, height = 6)


```









```{r}
library(dplyr)
library(ggplot2)

# Calculate the mean abundance for each combination of cell stage and strain
mean_data <- combined_data %>%
  group_by(Cell_Stage, Strain) %>%
  summarise(mean_abundance = mean(`lin 41 mRNA molecules`)) %>%
  arrange(Strain)  # Arrange the data by strain for easier segmentation

# Create a line plot
p <- ggplot(mean_data, aes(x = Cell_Stage, y = mean_abundance, color = Strain)) +
  geom_line() +
  geom_point() +
  # Add horizontal segments to connect points of the same strain
  geom_segment(data = mean_data %>% group_by(Strain) %>% slice(1) %>% filter(!is.na(lead(Cell_Stage))),
               aes(x = Cell_Stage, xend = lead(Cell_Stage), y = mean_abundance, yend = mean_abundance, color = Strain)) +
  labs(title = "Mean Abundance of lin-41 mRNA per Cell Stage",
       x = "Cell Stage",
       y = "Mean lin-41 mRNA abundance",
       color = "Strain") +
  theme_minimal() +
  theme(legend.position = "right")

# Display the plot
print(p)

```





```{r}

#To ask rate of decay

# Perform ANOVA
anova_result <- aov(`lin 41 mRNA molecules` ~ Cell_Stage * Strain, data = combined_data)

# Display ANOVA summary
summary(anova_result)

# Perform Tukey's post hoc test
tukey_result <- TukeyHSD(anova_result)

# Display Tukey's post hoc test results
print(tukey_result)


```



```{r}
# Perform ANOVA
anova_result <- aov(`lin 41 mRNA molecules` ~ subdirectory, data = combined_data)

# Display ANOVA summary
summary(anova_result)

# Perform Tukey's post hoc test
tukey_result <- TukeyHSD(anova_result)

# Display Tukey's post hoc test results
print(tukey_result)

```


The ANOVA results indicate that there are significant differences among the subdirectories (p < 0.001). Now, let's interpret the Tukey's post hoc test results:

DG5329 and DG5398 both have significantly higher lin 41 mRNA molecules compared to N2 (p < 0.001).
DG5779 also has significantly higher lin 41 mRNA molecules compared to N2 (p = 0.006).
DG5329 and DG5398 have significantly higher lin 41 mRNA molecules compared to DG5410 (p < 0.001).
DG5779 has significantly higher lin 41 mRNA molecules compared to DG5410 (p = 0.006).
DG5779 and DG5398 have a significant difference in lin 41 mRNA molecules (p = 0.017).



