---
title: "smFISH quantification report `r format(Sys.Date(), '%Y-%m-%d')`"
author: "Naly Torres"
date: "1/30/2024"
output: html_document
---


```{r echo=FALSE, warning=FALSE} 

library(png)
library(knitr)
library(ggplot2)
library(tidyr)
library(dplyr)

#install.packages("ggplot2")
library(ggsignif)
#install.packages("scales")
library(svglite)
library(ggridges)
library(scales)


```


Violin Plots

```{r}

# Parent directory containing CSV files for all conditions
parent_directory <- "/Users/nalytorres/Library/CloudStorage/OneDrive-Colostate/01_bigfish/02_RNAi-treatments-csv"


# Function to read CSV, replace Image ID with CSV name, and modify column names
read_modify_csv <- function(file_path, subdirectory) {
  my_df <- read.csv(file_path)
  my_df$Image_ID <- basename(file_path)
  my_df$subdirectory <- subdirectory
  my_df <- my_df[, colnames(my_df) != "X"]
  current_colnames <- colnames(my_df)
  new_colnames <- gsub("\\.", " ", current_colnames)
  colnames(my_df) <- new_colnames
  return(my_df)
}

# List all subdirectories (conditions) in the parent directory
subdirectories <- list.dirs(parent_directory, full.names = FALSE, recursive = FALSE)

# Initialize an empty list to store data frames
all_data <- list()

# Loop through subdirectories (conditions)
for (subdirectory in subdirectories) {
  # List all CSV files in the current subdirectory
  csv_files <- list.files(path = file.path(parent_directory, subdirectory), pattern = "*.csv", full.names = TRUE)
  
  # Loop through CSV files, read data, replace Image ID, and modify column names
  for (file_path in csv_files) {
    current_data <- read_modify_csv(file_path, subdirectory)
    all_data[[file_path]] <- current_data
  }
}

# Combine all data frames into one data frame
combined_data <- bind_rows(all_data)

combined_data
```



```{r}
#Horizontal

# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi", "DHC-1_RNAi")  # 

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Specify color for all levels
fill_colors <- c("L4440_RNAi" = "hotpink", "KLP-9_RNAi" = "lightgreen", "BMK-1_RNAi" = "lightgreen", "KLP-18_RNAi" = "lightgreen", "KLP-19_RNAi" = "lightgreen","UNC-116_RNAi" = "lightgreen", "NMY-1_RNAi" = "skyblue", "NMY-2_RNAi" = "skyblue", "DHC-1_RNAi" = "orange")

# Create a combined violin and boxplot
p <- ggplot(combined_data, aes(x = subdirectory, y = `erm 1 mRNA molecules`, fill = subdirectory)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 0.75), outlier.shape = NA) +
  geom_violin(alpha = 0.4, color = "white", size = 0.5, position = position_dodge(width = 0.75)) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 1) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knock down",
       x = "RNAi treatments",
       y = "erm-1 mRNA abundance \n (total molecules per embryo)",
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"),
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_x_discrete(labels = c("L4440_RNAi" = "Control",
                              "BMK-1_RNAi" = "KLP-14 ",
                              "KLP-18_RNAi" = "KLP-18",
                              "KLP-19_RNAi" = "KLP-19",
                              "KLP-9_RNAi" = "KLP-9  ",
                              "NMY-1_RNAi" = "NMY-1",
                              "UNC-116_RNAi" = "UNC-116",
                              "NMY-2_RNAi" = "NMY-2",
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(ylim = c(0, 7500)) +
  guides(fill = guide_legend(nrow = 2))  # Limit the legend to two rows

# Display the plot
print(p)

# Save the plot as SVG with specified output path
#ggsave("/Users/nalytorres/Documents/01_bigfish/RNAi-treatments-violin/RNAi-treatments-horizontal-plot.svg", p, width = 10, height = 6)



```

```{r}
# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi", "DHC-1_RNAi")

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Specify color for all levels
fill_colors <- c("L4440_RNAi" = "hotpink", "KLP-9_RNAi" = "lightgreen", "BMK-1_RNAi" = "lightgreen", "KLP-18_RNAi" = "lightgreen", "KLP-19_RNAi" = "lightgreen",
                 "UNC-116_RNAi" = "lightgreen", "NMY-1_RNAi" = "skyblue", "NMY-2_RNAi" = "skyblue", "DHC-1_RNAi" = "orange")

# Create a horizontal bar plot
p <- ggplot(combined_data, aes(y = subdirectory, x = `erm 1 mRNA molecules`, fill = subdirectory)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.75)) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knockdown",
       y = "RNAi treatments", 
       x = "erm-1 mRNA abundance \n (total molecules per embryo)", 
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors, 
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"), 
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_y_discrete(labels = c("L4440_RNAi" = "Control", 
                              "BMK-1_RNAi" = "KLP-14", 
                              "KLP-18_RNAi" = "KLP-18", 
                              "KLP-19_RNAi" = "KLP-19", 
                              "KLP-9_RNAi" = "KLP-9", 
                              "NMY-1_RNAi" = "NMY-1", 
                              "UNC-116_RNAi" = "UNC-116", 
                              "NMY-2_RNAi" = "NMY-2", 
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(size = 12), 
        axis.text.x = element_text(size = 12), 
        legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(color = "black"), 
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"), 
        axis.title = element_text(size = 13, face = "bold"), 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  coord_flip()  # Flip the axes

# Display the plot
print(p)

```


```{r}
# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi", "DHC-1_RNAi")

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Specify color for all levels
fill_colors <- c("L4440_RNAi" = "hotpink", "KLP-9_RNAi" = "purple", "BMK-1_RNAi" = "purple", "KLP-18_RNAi" = "purple", "KLP-19_RNAi" = "purple",
                 "UNC-116_RNAi" = "purple", "NMY-1_RNAi" = "skyblue", "NMY-2_RNAi" = "skyblue", "DHC-1_RNAi" = "orange")

# Create a boxplot with individual points
p <- ggplot(combined_data, aes(y = subdirectory, x = `erm 1 mRNA molecules`, fill = subdirectory)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Create boxplots without showing outliers
  geom_jitter(position = position_jitter(height = 0.2, width = 0), size = 2, alpha = 0.8, color = "black") +  # Add individual data points with jitter for clarity
  labs(title = "erm-1 mRNA abundance in embryos after motor knockdown",
       y = "RNAi treatments", 
       x = "erm-1 mRNA abundance \n (molecules per embryo)", 
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors, 
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"), 
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_y_discrete(labels = c("L4440_RNAi" = "Control", 
                              "BMK-1_RNAi" = "KLP-14", 
                              "KLP-18_RNAi" = "KLP-18", 
                              "KLP-19_RNAi" = "KLP-19", 
                              "KLP-9_RNAi" = "KLP-9", 
                              "NMY-1_RNAi" = "NMY-1", 
                              "UNC-116_RNAi" = "UNC-116", 
                              "NMY-2_RNAi" = "NMY-2", 
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(size = 12), 
        axis.text.x = element_text(size = 12), 
        legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(color = "black"), 
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"), 
        axis.title = element_text(size = 13, face = "bold"), 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 12)) +
  coord_flip()  # Flip the axes

# Display the plot
print(p)

```


```{r}
# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi", "DHC-1_RNAi")
#custom_order <- rev(custom_order)
# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Specify color for all levels
fill_colors <- c("L4440_RNAi" = "hotpink", "KLP-9_RNAi" = "lightgreen", "BMK-1_RNAi" = "lightgreen", "KLP-18_RNAi" = "lightgreen", "KLP-19_RNAi" = "lightgreen",
                 "UNC-116_RNAi" = "lightgreen", "NMY-1_RNAi" = "skyblue", "NMY-2_RNAi" = "skyblue", "DHC-1_RNAi" = "orange")

# Create a boxplot with individual points
p <- ggplot(combined_data, aes(x =subdirectory, y = `erm 1 mRNA molecules`, fill = subdirectory)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.6) +  # Create boxplots without showing outliers
  geom_jitter(position = position_jitter(width = 0.2, height = 0), size = 2, alpha = 0.8, color = "black") +  # Add individual data points with jitter for clarity
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knockdown",
       x = "Total erm-1 mRNA abundance \n (molecules per embryo)", 
       y = "RNAi treatments", 
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors, 
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"), 
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_x_discrete(labels = c("L4440_RNAi" = "Control", 
                              "BMK-1_RNAi" = "KLP-14", 
                              "KLP-18_RNAi" = "KLP-18", 
                              "KLP-19_RNAi" = "KLP-19", 
                              "KLP-9_RNAi" = "KLP-9", 
                              "NMY-1_RNAi" = "NMY-1", 
                              "UNC-116_RNAi" = "UNC-116", 
                              "NMY-2_RNAi" = "NMY-2", 
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1),  # Tilt x-axis labels for readability
        axis.text.y = element_text(size = 12), 
        legend.position = "right", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(), 
        axis.line = element_line(color = "black"), 
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"), 
        axis.title = element_text(size = 13, face = "bold"), 
        legend.title = element_text(size = 14), 
        legend.text = element_text(size = 12))

# Display the plot
print(p)

```




```{r}
#Horizontal - kinesin phenotypes

# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi")  

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Specify color for all levels
fill_colors <- c("L4440_RNAi" = "hotpink", "KLP-9_RNAi" = "lightgreen", "BMK-1_RNAi" = "lightgreen")

# Create a combined violin and boxplot
p <- ggplot(combined_data, aes(x = subdirectory, y = `erm 1 mRNA molecules`, fill = subdirectory)) +
  geom_violin(alpha = 0.4, color = "white", size = 0.5, position = position_dodge(width = 0.75)) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.75), outlier.shape = NA) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 1) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knock down",
       x = "RNAi treatments",
       y = "erm-1 mRNA abundance \n (total molecules per embryo)",
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("L4440_RNAi", "KLP-9_RNAi"),
                    labels = c("Control", "Kinesin")) +
  scale_x_discrete(labels = c("L4440_RNAi" = "Control",
                              "BMK-1_RNAi" = "BMK-1",
                              "KLP-9_RNAi" = "KLP-9")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(ylim = c(0, 7500)) +
  guides(fill = guide_legend(nrow = 2))  # Limit the legend to two rows

# Display the plot
print(p)


# Save the plot as SVG with specified output path
#ggsave("/Users/nalytorres/Documents/01_bigfish/RNAi-treatments-violin/RNAi-treatments-horizontal-kinesin-phenotypes-plot.svg", p, width = 10, height = 6)



```


```{r}
#Horizontal - myosin phenotypes

# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "NMY-1_RNAi", "NMY-2_RNAi")  

# Convert subdirectory to a factor with custom order
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = custom_order)

# Specify color for all levels
fill_colors <- c("L4440_RNAi" = "hotpink", "NMY-1_RNAi" = "lightblue", "NMY-2_RNAi" = "lightblue")

# Create a combined violin and boxplot
p <- ggplot(combined_data, aes(x = subdirectory, y = `erm 1 mRNA molecules`, fill = subdirectory)) +
  geom_violin(alpha = 0.4, color = "white", size = 0.5, position = position_dodge(width = 0.75)) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.75), outlier.shape = NA) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 1) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knock down",
       x = "RNAi treatments",
       y = "erm-1 mRNA abundance \n (total molecules per embryo)",
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("L4440_RNAi", "NMY-2_RNAi"),
                    labels = c("Control", "Myosin")) +
  scale_x_discrete(labels = c("L4440_RNAi" = "Control",
                              "NMY-2_RNAi" = "NMY-2",
                              "NMY-1_RNAi" = "NMY-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(ylim = c(0, 7500)) +
  guides(fill = guide_legend(nrow = 2))  # Limit the legend to two rows

# Display the plot
print(p)


# Save the plot as SVG with specified output path
ggsave("/Users/nalytorres/Documents/01_bigfish/RNAi-treatments-violin/RNAi-treatments-horizontal-myosin-phenotypes1-plot.svg", p, width = 10, height = 6)


```



```{r}
# Vertical

# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi")  # 
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = rev(custom_order))

# Create a combined violin and boxplot
p <- ggplot(combined_data, aes(x = `erm 1 mRNA molecules`, y = subdirectory, fill = subdirectory)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 0.75), outlier.shape = NA) +
  geom_violin(alpha = 0.4, color = "white", size = 0.5, position = position_dodge(width = 0.75)) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 1) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knock down",
       x = "erm-1 mRNA abundance \n (total molecules per embryo)",
       y = "RNAi treatments",
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"),
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_y_discrete(labels = c("L4440_RNAi" = "Control",
                              "BMK-1_RNAi" = "KLP-14 ",
                              "KLP-18_RNAi" = "KLP-18",
                              "KLP-19_RNAi" = "KLP-19",
                              "KLP-9_RNAi" = "KLP-9  ",
                              "NMY-1_RNAi" = "NMY-1",
                              "UNC-116_RNAi" = "UNC-116",
                              "NMY-2_RNAi" = "NMY-2",
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(angle = 0, hjust = 1),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(xlim = c(0, 7500)) +
  guides(fill = guide_legend(nrow = 3))  # Limit the legend to two rows

# Display the plot
print(p)

# Save the plot as SVG with specified output path
ggsave("//Users/nalytorres/Library/CloudStorage/OneDrive-Colostate/01_bigfish/02_RNAi-treatments-csv/RNAi-treatments-plot.svg", p, width = 10, height = 6)




```
```{r}
# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi")
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = rev(custom_order))

# Create a combined violin and boxplot
p <- ggplot(combined_data, aes(x = `erm 1 mRNA molecules`, y = subdirectory, fill = subdirectory)) +
  geom_boxplot(width = 0.25, position = position_dodge(width = 0.75), outlier.shape = NA) +
  geom_violin(alpha = 0.4, color = "white", size = 0.5, position = position_dodge(width = 0.75)) +
  geom_jitter(height = 0, width = 0.2, color = "black", size = 1) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knock down",
       x = "erm-1 mRNA abundance \n (total molecules per embryo)",
       y = NULL,  # Removed y-axis label
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"),
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_y_discrete(labels = c("L4440_RNAi" = "Control",
                              "BMK-1_RNAi" = "KLP-14 ",
                              "KLP-18_RNAi" = "KLP-18",
                              "KLP-19_RNAi" = "KLP-19",
                              "KLP-9_RNAi" = "KLP-9  ",
                              "NMY-1_RNAi" = "NMY-1",
                              "UNC-116_RNAi" = "UNC-116",
                              "NMY-2_RNAi" = "NMY-2",
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size = 13, face = "bold", margin = margin(r = 20)),  # Adjusted y-axis label position
        axis.title.x = element_text(size = 13, face = "bold"),  # Added x-axis label formatting
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_flip() +  # Flip the plot vertically
  coord_cartesian(xlim = c(0, 7500)) +
  guides(fill = guide_legend(nrow = 3))  # Limit the legend to two rows

# Display the plot
print(p)

# Save the plot as SVG with specified output path
ggsave("//Users/nalytorres/Library/CloudStorage/OneDrive-Colostate/01_bigfish/02_RNAi-treatments-csv/RNAi-treatments-plot.svg", p, width = 10, height = 6)

```




```{r}
# Load necessary libraries


# Define the custom order for subdirectories
custom_order <- c("L4440_RNAi", "KLP-9_RNAi", "BMK-1_RNAi", "KLP-18_RNAi", "KLP-19_RNAi", "UNC-116_RNAi", "NMY-1_RNAi", "NMY-2_RNAi")  # 
combined_data$subdirectory <- factor(combined_data$subdirectory, levels = rev(custom_order))

# Create the ridgeline plot
p <- ggplot(combined_data, aes(x = `erm 1 mRNA molecules`, y = subdirectory, fill = subdirectory)) +
  geom_density_ridges(scale = 0.9, rel_min_height = 0.01, alpha = 0.7) +
  labs(title = "Total erm-1 mRNA abundance in embryos after motor knock down",
       x = "erm-1 mRNA abundance \n (total molecules per embryo)",
       y = "RNAi treatments",
       fill = "Motor Class") +  # Set the legend title here
  scale_fill_manual(values = fill_colors,
                    breaks = c("L4440_RNAi", "KLP-9_RNAi", "NMY-2_RNAi", "DHC-1_RNAi"),
                    labels = c("Control", "Kinesin", "Myosin", "Dynein")) +
  scale_y_discrete(labels = c("L4440_RNAi" = "Control",
                              "BMK-1_RNAi" = "KLP-14 ",
                              "KLP-18_RNAi" = "KLP-18",
                              "KLP-19_RNAi" = "KLP-19",
                              "KLP-9_RNAi" = "KLP-9  ",
                              "NMY-1_RNAi" = "NMY-1",
                              "UNC-116_RNAi" = "UNC-116",
                              "NMY-2_RNAi" = "NMY-2",
                              "DHC-1_RNAi" = "DHC-1")) +
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(angle = 45, hjust = 1),
        legend.position = "right",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
        plot.subtitle = element_text(size = 14),
        axis.title = element_text(size = 13, face = "bold"),
        axis.text = element_text(size = 12),
        axis.ticks = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12)) +
  coord_cartesian(xlim = c(0, 7500)) +
  guides(fill = guide_legend(nrow = 3))  # Limit the legend to two rows

# Display the plot
print(p)

```



```{r}
# Shapiro-Wilk test for normality
shapiro.test(combined_data$`erm 1 mRNA molecules`)
```
The Shapiro-Wilk normality test is used to assess whether a given dataset is normally distributed. Here’s a brief interpretation of your test results:

W statistic (0.78378): This is the test statistic for the Shapiro-Wilk test. It ranges from 0 to 1, with values closer to 1 indicating that the data is more likely to be normally distributed.
p-value (< 2.2e-16): This is the probability of observing the test results under the null hypothesis that the data is normally distributed. A very small p-value (typically less than 0.05) suggests that the data is not normally distributed.
In your case, the W statistic is 0.78378, and the p-value is less than 2.2e-16, which is extremely small. This indicates that there is strong evidence against the null hypothesis, suggesting that the erm 1 mRNA molecules data is not normally distributed.

Would you like more information on how to proceed given these results?




```{r}

# Kruskal-Wallis test
kruskal.test(`erm 1 mRNA molecules` ~ subdirectory, data = combined_data)


```

The Kruskal-Wallis rank sum test is a non-parametric method used to determine if there are statistically significant differences between the medians of three or more independent groups. Here’s how to interpret the results you provided:

Interpretation
Kruskal-Wallis chi-squared (39.678): This is the test statistic for the Kruskal-Wallis test. Higher values indicate a greater difference between groups.
Degrees of freedom (df = 7): This is the number of groups minus one. In your case, it suggests that there are 8 groups (7 + 1).
p-value (1.451e-06): This is the probability of observing the test results under the null hypothesis that all group medians are equal. A very small p-value (typically less than 0.05) indicates that at least one group median is significantly different from the others.
Since your p-value is 1.451e-06, which is much smaller than 0.05, you can reject the null hypothesis. This means there is strong evidence that at least one of the groups has a different median erm 1 mRNA molecules level compared to the others.


```{r}
# Perform the Dunn test
dunn.test(combined_data$`erm 1 mRNA molecules`, combined_data$subdirectory, method = "bonferroni")
```



```{r}
# Boxplot
boxplot(`erm 1 mRNA molecules` ~ subdirectory, data = combined_data, main = "erm 1 mRNA molecules by Subdirectory", xlab = "Subdirectory", ylab = "erm 1 mRNA molecules")

```




```{r}
# Perform ANOVA -> can't do this because the data is not normally distributed
anova_result <- aov(`erm 1 mRNA molecules` ~ subdirectory, data = combined_data)

# Display ANOVA summary
summary(anova_result)

# Perform Tukey's post hoc test
tukey_result <- TukeyHSD(anova_result)

# Sort Tukey's test results alphabetically by comparisons
sorted_tukey_result <- tukey_result$subdirectory[order(row.names(tukey_result$subdirectory)), ]

# Display sorted Tukey's post hoc test results
# Format p-values to avoid scientific notation
sorted_tukey_result <- data.frame(sorted_tukey_result)
sorted_tukey_result$p.adj <- format(sorted_tukey_result$p.adj, scientific = FALSE)

sorted_tukey_result


```





